Wait-free Queues with Polylogarithmic Step Complexity

### 2022-09-29 thesis/thesis.tex
update figures

We show $EST_{n\nf{.left}}^{t^R}= ops(n\nf{.left.blocks[0..n.left.head$_{309}-1$]}) \subseteq ops(n.blocks_{\ref{cas}}) = ops(n\nf{.blocks[0..}n.head_{\ref{cas}}\nf{]})$.

Line \ref{cas} stores a block \nf{new} in $n$ that has $\nf{end\sub{left}}=n.left.head_{\ref{lastLine}}-1$. Therefore by Definition \ref{def::subblock}, after the successful \nf{CAS} in Line \ref{cas} we know all blocks in \nf{n.left.blocks[1..$n.left.head_{\ref{lastLine}}-1$]} are subblocks of \nf{n.blocks[1..$n.head_{\ref{readHead}}$]}. Because of Lemma \ref{nonDecreasingHead} we have $n.left.head_{309}-1<n.left.head_{\ref{lastLine}}-1$ and $n.head_{\ref{readHead}}<n.head_{\ref{cas}}$. From Observation \ref{lem::blocksOrder} the claim follows.  The proof for the right child is the same.
\end{proof}
\begin{figure}[hbt]
  \center\includegraphics[width=5in]{pics/trueRefresh}
  \caption{New established operations of the left child are in the new block. (TO UPDATE)}
\end{figure}

@@

\begin{figure}[hbt]  
  \center\includegraphics[width=7in]{pics/timeLineSuper1.png}
  \caption{Time relations between \nf{R\sub{n},R\sub{p},R',R"}}
  \label{fig::timeLine1}
\end{figure}

@@

Let the first successful \nf{p.Refresh()} be \nf{Rp1} and the second next successful \nf{p.Refresh()} be \nf{Rp2}. If \nf{Rp1} reads \nf{x} in Line \ref{readHead}, then \nf{Rp2} has to read \nf{x+1} in Line \ref{readHead} (iduced from \ref{lem::headProgress}, \ref{nonDecreasingHead}). See the timeline in Figure \ref{fig::timeLine2} for two consecutive successful \nf{Refresh()} instances \nf{Rp1, Rp2} on \nf{p}.

\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/timeLineSuper2.png}
  \caption{Rp2 reads p.head after $t_{321}^{Rp1}$, which is after $t_{321}^{Rn}$. Rp2 has to help increment n.head and set b.super.}
  \label{fig::timeLine2}
\end{figure}

\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/timeLineSuper3.png}
  \caption{ The second Refresh on \nf{p} contains \nf{b} and reads $\nf{n.head}>\nf{i}$.}
  \label{fig::timeLine3}
\end{figure}

@@

\begin{figure}[hbt]  
  \center\includegraphics[width=2in]{pics/computeResponseDetail.png}\caption{The position of $D_i(root,b)$.}
\label{computeResponseDetail}
\end{figure}

@@

So the doubling search reaches \nf{start} such that the \nf{root.blocks[start].sum\sub{enq}} is less than $e$ in $O \big(\log(\nf{root.blocks[$b$].size}+\nf{root.blocks[$end$].size})\big)$ steps. See Figure \ref{fig::doubling}. After Line \ref{dsearchEnd}, the binary search that finds \nf{b} also takes $O\big(\log(\nf{root.blocks[$b$].size}+\nf{root.blocks[$end$].size})\big)$. Next, \nf{i} is computed via the definition of \nf{sum\sub{enq}} in constant time (Line \ref{DSearchComputei}). So the claim is proved.
\end{proof}
\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/doubling.png}
  \caption{Distance relations between \nf{start}$,b,end$. \nf{start} is shown with $c$.}
  \label{fig::doubling}
\end{figure}

>>>>

We show 
\begin{align*}
EST_{n\nf{.left}}^{t^R}&= ops(n\nf{.left.blocks[0..n.left.head$_{309}-1$]}) \\
&\subseteq ops(n.blocks_{\ref{cas}}) = ops(n\nf{.blocks[0..}n.head_{\ref{cas}}\nf{]}).
\end{align*}

Line \ref{cas} stores a block \nf{new} in $n$ that has $\nf{end\sub{left}}=n.left.head_{\ref{lastLine}}-1$. Therefore by Definition \ref{def::subblock}, after the successful \nf{CAS} in Line \ref{cas} we know all blocks in \nf{$n$.left.blocks[$1\cdots n.left.head_{\ref{lastLine}}-1$]} are subblocks of \nf{n.blocks[$1\cdots n.head_{\ref{readHead}}$]}. Because of Lemma \ref{nonDecreasingHead} we have $n.left.head_{309}-1<n.left.head_{\ref{lastLine}}-1$ and $n.head_{\ref{readHead}}<n.head_{\ref{cas}}$. From Observation \ref{lem::blocksOrder} the claim follows.  The proof for the right child is the same.
\end{proof}

\begin{figure}[hbt]
  \center\includegraphics[width=5in]{pics/trueRefresh}
  \caption{TODO :: Is this picture good or should I remove it? New established operations of the left child are in the new block.(TO UPDATE)}
\end{figure}

@@

\begin{figure}[hbt]  
  \center\includegraphics[width=5.5in,height=0.65in]{pics/timeLineSuper1.png}
  \caption{Time relations between \nf{R\sub{n},R\sub{p},R',R"}}
  \label{fig::timeLine1}
\end{figure}

@@

Let the first successful \nf{p.Refresh()} be $R_{p1}$ and the second next successful \nf{p.Refresh()} be $R_{p2}$. If $R_{p1}$ reads \nf{x} in Line \ref{readHead}, then $R_{p2}$ has to read \nf{x+1} in Line \ref{readHead} (induced from \ref{lem::headProgress}, \ref{nonDecreasingHead}). See the timeline in Figure \ref{fig::timeLine2} for two consecutive successful \nf{Refresh()} instances $R_{p1}$,$R_{p2}$  on ${p}$.

\begin{figure}[hbt]  
  \center\includegraphics[width=6in,height=1.25in]{pics/timeLineSuper2.png}
  \caption{Rp2 reads p.head after $t_{321}^{Rp1}$, which is after $t_{321}^{Rn}$. Rp2 has to help increment n.head and set b.super.}
  \label{fig::timeLine2}
\end{figure}

$$t_{\ref{cas}}^{R_n}<t_{\ref{cas}}^{R_p}<t_{\ref{cas}}^{R_p'}<t_{314}^{R_p'}$$
$t_{\ref{cas}}^{R_n}<t_{\ref{cas}}$ because $R_p$ is the first successful $p.Refresh()$ after $R_n$. $t_{\ref{cas}}^{R_p}<t_{\ref{cas}}^{R_p'}$ because of the time windows between two consecutive successful \nf{Refresh()} instances on a node. $t_{\ref{cas}}^{R_p'}<t_{314}^{R_p'}$ because it is sequential in the code. The second Refresh on \nf{p} contains \nf{b} and reads $\nf{n.head}>\nf{i}$.

@@

\begin{figure}[hbt]  
  \center\includegraphics[width=4in, height=2.5in]{pics/computeResponseDetail.png}\caption{The position of $D_i(root,b)$.}
\label{computeResponseDetail}
\end{figure}

@@

So the doubling search reaches \nf{start} such that the \nf{root.blocks[start].sum\sub{enq}} is less than $e$ in $O \big(\log(\nf{root.blocks[$b$].size}+\nf{root.blocks[$end$].size})\big)$ steps. See Figure \ref{fig::doubling}. After Line \ref{dsearchEnd}, the binary search that finds \nf{b} also takes $O\big(\log(\nf{root.blocks[$b$].size}+\nf{root.blocks[$end$].size})\big)$. Next, \nf{i} is computed via the definition of \nf{sum\sub{enq}} in constant time (Line \ref{DSearchComputei}). So the claim is proved.
\end{proof}
\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/doubling.png}
  \caption{Distance relations between \nf{start}$,b,end$.}
  \label{fig::doubling}
\end{figure}
